The SMT Competition 2015--2018

### 2018-09-21 2018/report/SMT-COMP.tex
Add JSAT Call for Papers
First draft of participants section

@@

@@

\begin{abstract}
TODO
\end{abstract}

@@

@@

@@

@@

>>>>

\usepackage{pifont}
\usepackage{comment}

@@

\specialcomment{tjark}{\begingroup\footnotesize\color{red}}{\color{black}\endgroup}
\newcommand{\TODO}[0]{\textcolor{red}{TODO}}

\renewcommand{\mark}[0]{\ding{51}}

% TODO
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
\TODO
\end{abstract}

@@

%% Expected content according to the JSAT Call for Papers:

%% Full articles by competition organizers reporting on SAT 2018
%% competitions and evaluation. The articles should describe the
%% competition, its criteria, why it is interesting to the SAT
%% research community, execution environment used, analysis of the
%% results (including how they compare to previous instantiations, if
%% appropriate), and give a summary of the main technical
%% contributions to the field, as well as discussions on lessons
%% learned and suggestions for improvements for future competitions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{tjark}
Changes 2015:
\begin{itemize}
\item All solvers will be run with $n=4$ cores available.  We will
  recognize both best sequential and best parallel performance in all
  main track divisions, using different CPU time limits.
\item The division scoring will take time spent on unsolved benchmarks
  into account.  Previous competitions only considered time spent on
  solved benchmarks.  However, the number of solved benchmarks still
  takes precedence over run-time.
\item In addition to recognizing the best solver in each division, we
  will recognize solvers that perform best according to
  competition-wide criteria, emphasizing breadth of supported logics.
  These criteria will be a refinement of the criteria used for the
  FLoC Olympic Games medals in~2014.
\item We plan to include new divisions for floating-point arithmetic.
  These should be considered experimental in~2015, and will not be
  included in the competition-wide ranking.
\item We plan to run solvers on all eligible benchmarks.  Previous
  competitions used a proper subset of eligible benchmarks, causing
  competition results to be affected by a random selection process.
\end{itemize}

Changes 2016:
\begin{itemize}
\item Benchmarks will use (a subset of) version~2.5 of the SMT-LIB
  language.  \emph{Rationale:} This is the latest version of the
  SMT-LIB language.  It was released on 2015-06-28, and is largely
  backwards-compatible to version~2.0, which was used for SMT-COMP
  2015.
\item Solver output may affect the score even when the solver does not
  terminate within the time limit.  (In 2015, main track solver output
  was ignored if the solver subsequently timed out.)  Solvers should
  take care not to accidentally produce output that contains
  \texttt{sat} or \texttt{unsat} even when they are killed.
  \emph{Rationale:} Users are likely to trust solver responses even
  when the solver continues to run for some time.
\item Divisions for floating-point arithmetic are no longer
  experimental, and will be considered competitive if the necessary
  requirements (see Section~\ref{sec:scoring}) are met.
  \emph{Rationale:} Floating-point divisions were experimental in
  2015.  By now, their definition is sufficiently stable, and they are
  supported by several solvers.
\item Division scores will be based on a weighted sum of scores for
  benchmark families.  \emph{Rationale:} For some years now, SMT-COMPs
  have had sufficient computational resources to evaluate all solver
  entrants on all eligible benchmarks.  As a side-effect, the
  weighting of benchmark families that was achieved in early SMT-COMPs
  through a careful selection of benchmarks (based on benchmark
  difficulty and category) was lost.  Since there are vast differences
  in size between benchmark families, we believe that a weighting that
  de-emphasizes large benchmark families (see
  Section~\ref{sec:division-scoring}) will lead to more meaningful
  competition results.
\item The unsat-core track that was introduced in SMT-COMP 2012, but
  discontinued in 2013-2015 (partly because of the competition's move
  to StarExec), is back.  It will be experimental in 2016; its results
  will be reported, but no official winners will be announced.
  \emph{Rationale:} Unsat cores are important in many applications of
  SMT solving.  The track was discontinued primarily for
  infrastructure and resource reasons.  We are pleased that we can
  finally offer it again, but we ask for your understanding in case of
  mishaps while we port the required tools.
\item Best industrial performance will no longer be recognized
  separately.  \emph{Rationale:} While there is agreement in the SMT
  community to emphasize problems that come from real applications,
  industrial performance largely coincided with overall performance in
  SMT-COMP 2015.  We consider the effort to determine and report it
  separately no longer justified.
\end{itemize}

Changes 2017:
\begin{itemize}
\item Submission of accompanying information for solvers is via a web
  form (rather than by email to the organizers).  \emph{Rationale:}
  Email submissions were often incomplete, leading to ambiguity and
  further inquiries.  The form provides a more structured way of
  submitting information.
\item Benchmarks will use (a subset of) version~2.6 of the SMT-LIB
  language.  \emph{Rationale:} This is the latest version of the
  SMT-LIB language.  It is largely backwards-compatible with earlier
  versions (in particular~2.0 and~2.5).
\item Benchmarks that use partial bit-vector functions, such as
  \texttt{bvudiv}, or underspecified floating-point functions, such as
  \texttt{fp.min}, will be eligible for the competition again.
  \emph{Rationale:} A conclusion on the semantics of these functions
  has finally been reached, and the status of affected SMT-LIB
  benchmarks has been updated accordingly.
\item Benchmarks with unknown status will be eligible for the
  competition's main track.  In 2016, solver performance on benchmarks
  with unknown status was evaluated but reported separately; in 2017,
  it will directly affect the competition results.  \emph{Rationale:}
  Using only benchmarks with known status (i.e., benchmarks that have
  been solved before) to determine the competition results unjustly
  favors imitation of existing solvers over true innovation.
\item In the competition-wide scoring, the (per-division) penalty for
  erroneous results is changed from the number of errors to a fixed
  constant (multiplied by the weight of the respective division).
  \emph{Rationale:} The weight already takes the number of benchmarks
  in the division into account; multiplying it by the number of errors
  overemphasizes large divisions.
\item The unsat-core track, which was re-introduced in 2016, is no
  longer experimental.  Its divisions will be considered competitive
  if the necessary requirements (see Section~\ref{sec:scoring}) are
  met.  \emph{Rationale:} Rules and tool support for the unsat-core
  track proved sufficiently stable in 2016.
\item The competition will feature experimental divisions for
  benchmarks that use algebraic datatypes.  \emph{Rationale:}
  Algebraic datatypes are specified in the draft SMT-LIB~2.6 standard,
  which is expected to become official in the near future, and are
  already supported by some solvers.  Benchmarks that use datatypes
  have recently been added to SMT-LIB.
\end{itemize}

Changes 2018:
\begin{itemize}
\item The time limit per solver/benchmark pair is anticipated to be at
  most 20 minutes in the main track, and at most 40 minutes in the
  other tracks.  \emph{Rationale:} Already in 2017, the time limit for
  the main track was reduced to 20 minutes (down from 40 minutes in
  earlier years) to cope with the inclusion of a large number of
  benchmarks with unknown status.
\item In the unsat-core track, the unsatisfiability check for the
  returned core will be based on a simple majority vote.  The case
  that a solver finds its own core satisfiable will no longer be
  treated specially.  \emph{Rationale:} This allows to simplify the
  post-processing of unsat-core track results considerably.
\item In 2017, the competition featured experimental divisions for
  benchmarks that use algebraic datatypes.  These divisions will be
  regular (non-experimental) in 2018. \emph{Rationale:} Algebraic
  datatypes are specified in the latest official release (Version~2.6)
  of the SMT-LIB standard.
\item The competition will feature an experimental divisions for
  benchmarks that use strings.  \emph{Rationale:} Corresponding
  theories and benchmarks are expected to be added to SMT-LIB in the
  near future.
\end{itemize}
\end{tjark}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Participants must upload their solvers to StarExec (see
Section~\ref{sec:starexec}) to enter them into the competition, and
additionally submit information about each competing solver---such as
the tracks and divisions into which the solver is being entered---to
the competition organizers.  Until 2016, this information was
collected by email.  Since 2017, a web form is being used for this
purpose.  This has greatly reduced the number of incomplete or
ambiguous submissions.

Table~\ref{table:participants} lists the solvers that participated
between~2015 and~2018, together with their submitter(s) and
institutional affiliations (which may have changed over time).  Note
that most solvers were developed by a group of people; the table does
not list all contributors but only submitters.
%
Participants are also encouraged to submit a short system description
of their solver, including a list of its authors and an explanation of
the SMT solving approach used.  System descriptions are published on
the SMT-COMP website~\cite{smtcomp-web}.  They were submitted for
about 60\,\% of all solver versions since 2015.

\begin{table}
  \caption{Participants (2015--2018)}
  \label{table:participants}
  \centering
  \begin{tabular}{|l|p{13cm}|} %TODO: fix width
    \hline
    Solver & Submitters (Affiliations)\\
    \hline
    ABC [\TODO]       & Valera Balabanov
                        (Mentor Graphics)\\
    Alt-Ergo          & Albin Coquereau
                        (CNRS and University of Paris-Sud)\\
    AProVE            & Carsten Fuhs
                        (University College London; Birkbeck, University of London)\\
    Boolector         & Mathias Preiner
                        (Johannes Kepler University)\\
    COLIBRI           & Fran\c{c}ois Bobot
                        (French Alternative Energies and Atomic Energy Commission)\\
    Ctrl-Ergo         & Mohamed Iguernlala
                        (CNRS and University of Paris-Sud)\\
    CVC3              & Kshitij Bansal
                        (New York University)\\
    CVC4              & Kshitij Bansal, Clark Barrett, Andres N\"otzli, Duligur Ibeling
                        (New York University; Stanford University)\\
    MapleSTP          & Jimmy Liang
                        (University of Waterloo)\\
    MinkeyRink        & Trevor Hansen
                        (University of Melbourne)\\
    OpenSMT2          & Matteo Marescotti, Antti Hyv\"arinen
                        (University of Lugano)\\
    ProB              & Sebastian Krings
                        (University of D\"usseldorf)\\
    Q3B               & Martin Jon\'a\v{s}
                        (Masaryk University)\\
    raSAT             & Tung Vu Xuan
                        (Japan Advanced Institute of Science and Technology)\\
    Redlog            & Haniel Barbosa
                        (Inria Nancy)\\
    SMT-RAT           & Florian Corzilius, Gereon Kremer
                        (RWTH Aachen University)\\
    SMTInterpol       & J\"urgen Christ
                        (University of Freiburg)\\
    SPASS-SATT        & Martin Bromberger
                        (Max Planck Institute for Informatics)\\
    STP               & Mate Soos, Trevor Hansen, Norbert Manthey
                        (National University of Singapore; University of Melbourne; Amazon Web Services)\\
    toysmt            & Masahiro Sakai
                        (\TODO)\\
    Vampire           & Giles Reger
                        (University of Manchester)\\
    veriT             & David D\'eharbe, Haniel Barbosa, Hans-J\"org Schurr
                        (Federal University of Rio Grande do Norte; Inria Nancy)\\
    XSat              & Martin Velez
                        (University of California, Davis)\\
    Yices2            & Bruno Dutertre, Dejan Jovanovi\'c
                        (SRI International)\\
    Z3                & Christoph Wintersteiger, Aleksandar Zelji\'c
                        (Microsoft Research; Uppsala University)\\
    \hline
  \end{tabular}  
\end{table}

Many of these solvers were submitted in multiple years.  Additionally,
to promote a wide comparison between tools, developers are allowed to
submit multiple versions of a solver, e.g., for different competition
tracks or with differently tuned internal components.
Table~\ref{table:participants-history} shows in more detail how many
versions of each solver were submitted in each year.

\begin{table}
  \caption{Number of versions submitted for each solver (by year).
    Numbers in square brackets indicate solver versions that were
    participating hors concours.}
  \label{table:participants-history}
  \centering
  \begin{tabular}{|l|r@{\,\,}r|r@{\,\,}r|r@{\,\,}r|r@{\,\,}r|}
    \hline
    Solver & \multicolumn{2}{c|}{2015} & \multicolumn{2}{c|}{2016} & \multicolumn{2}{c|}{2017} & \multicolumn{2}{c|}{2018} \\
    \hline
    ABC               &    &      &  2 &      &    &      &    &      \\
    Alt-Ergo          &    &      &    &      &    &      &  1 &      \\
    AProVE            &  1 &      &  1 &      &  1 &      &  1 &      \\
    Boolector         &  3 & [+1] &  2 &      &  2 &      &  2 &      \\
    COLIBRI           &    &      &    &      &  1 &      &  1 &      \\
    Ctrl-Ergo         &    &      &    &      &    &      &  1 &      \\
    CVC3              &  2 &      &    &      &    &      &    &      \\
    CVC4              &  4 &      &  1 & [+1] &  3 &      &  3 & [+1] \\
    MapleSTP          &    &      &  4 &      &    &      &    &      \\
    {[}MathSAT{]}     &    & [+2] &    & [+3] &    & [+3] &    & [+3] \\
    MinkeyRink        &    &      &  1 &      &  1 &      &  2 &      \\
    OpenSMT2          &  2 &      &  1 &      &  1 &      &    & [+1] \\
    ProB              &    &      &  1 &      &    &      &    &      \\
    Q3B               &    &      &  1 &      &  1 &      &  1 &      \\
    raSAT             &  1 &      &  2 &      &    &      &    &      \\
    Redlog            &    &      &    &      &  1 &      &    &      \\
    SMT-RAT           &  2 &      &  1 &      &  1 &      &  2 &      \\
    SMTInterpol       &  1 &      &  1 &      &  1 &      &  1 &      \\
    SPASS-SATT        &    &      &    &      &    &      &  1 &      \\
    STP               &  4 &      &  8 &      &  2 &      &  3 &      \\
    toysmt            &    &      &  1 &      &    &      &    &      \\
    Vampire           &    &      &  2 &      &  1 &      &  1 &      \\
    veriT             &  1 &      &  1 &      &  3 &      &  2 &      \\
    XSat              &    &      &    &      &  1 &      &    &      \\
    Yices2            &  3 &      &  2 &      &  2 &      &  3 &      \\
    Z3                &  2 & [+1] &    & [+1] &    & [+1] &    & [+1] \\
    \hline
    Total             & 26 & [+4] & 32 & [+5] & 22 & [+4] & 25 & [+6] \\
    \hline
  \end{tabular}
\end{table}

In addition to the solvers that were submitted to the competition by
their respective developers, the organizers included the most recent
stable versions of MathSAT and Z3 for comparison purposes.  Both
solvers are strong tools, but---except for two experimental versions
of Z3 in 2015---their development teams chose not to prepare
competition versions.  These solvers were therefore participating hors
concours.  Moreover, in 2015 a bug was discovered in the application
track version of Boolector and a fixed version was submitted after the
deadline; in 2016, the CVC4 team did not enter their solver into the
application track; and in 2018, an experimental version of CVC4 as
well as the OpenSMT2 solver were submitted after the deadline.  These
solvers were also participating hors concours.  In result tables, they
are listed with their name in square brackets (e.g., [MathSAT]).

Table~\ref{table:participation-by-track} summarizes these figures and
shows how many solver versions were submitted to each track of the
competition.  Note that solver versions may be entered into multiple
tracks.  Therefore, the total number of solver versions for each year
is typically less than the sum over all tracks.

By number of solver versions submitted, the competitions in 2015--2018
were the four largest in the history of SMT-COMP.  On average, 26
solver versions were submitted each year since 2015.  In contrast, the
competitions from 2005--2014 only received an average of 12
submissions~\cite{smtcomp-2014}.

\begin{table}
  \caption{Participation by track and year.  Numbers in square
    brackets indicate solver versions that were participating hors
    concours.}
  \label{table:participation-by-track}
  \centering
  \begin{tabular}{|l|r@{\,\,}r|r@{\,\,}r|r@{\,\,}r|r@{\,\,}r|}
    \hline
                      & \multicolumn{2}{c|}{2015} & \multicolumn{2}{c|}{2016} & \multicolumn{2}{c|}{2017} & \multicolumn{2}{c|}{2018} \\
    \hline
    Main track        & 21 &                [+2] & 25 & [+2] & 19 & [+2] & 20 & [+4] \\
    Application track & 10 &                [+3] &  8 & [+3] &  4 & [+2] &  4 & [+2] \\
    Unsat-core track  & \multicolumn{2}{c|}{---} &  1 & [+4] &  2 & [+2] &  3 & [+2] \\
    \hline
    Total             & 26 &                [+4] & 32 & [+5] & 22 & [+4] & 25 & [+6] \\
    \hline
  \end{tabular}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Local Variables:
% ispell-local-dictionary: "american"
% mode: LaTeX
% mode: flyspell
% LocalWords: concours hors satisfiability SMT Tjark
% End:
